---
output: github_document
---



```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CensoredMediator

<!-- badges: start -->
<!-- badges: end -->

The goal of CensoredMediator is analyze the mediation model with a censored mediator, described in:

"Wang J, Ning J, Shete S. Mediation analysis in a case-control study when the mediator is a censored variable. Stat Med 38(7):1213-1229, 3/2019. PMCID: PMC6467083" 

"Wang J, Ning J, Shete S. Mediation model with a categorical exposure and a censored mediator with application to a genetic study. PLoS One 16(10):e0257628, 2021. e-Pub 10/2021. PMCID: PMC8509986"

## Installation

You can install the development version of CensoredMediator like so:

``` r
library(devtools)
devtools::install_github("lilyxj91/CensoredMediator",force = TRUE)
```

## Example


```{r example}
library(CensoredMediator)
## basic example code
```

### Read the data

```{r}
### Read the data
data<-TestData
head(data)

# where 
# Y is the outcome;
# X is the initial variable;
# M is the censored mediator;
# delta is the censored status for M (1-observed; 0-censored);
# Z1 and Z2 are two covariates.  
# 
# Please arrange your data into this format, with the first four columns named as Y, X, M and delta; and then include other covariates from column five (Z1, Z2, Z3...).  
```


### Define the parameters
```{r}
## If the outcome is a continuous variable: 
## use the following parameters
# outcome.type<-"continuous"
# casecontrol<-FALSE
# prev<-0
# boot<-200  ## number of bootstrap samples
```

```{r}
## If the outcome is a binary variable but the study is not a case-control study: 
## use the following parameters
outcome.type<-"binary"
casecontrol<-FALSE
prev<-0
boot<-200  ## number of bootstrap samples
```

```{r}
## If the outcome is a binary variable and the study is a case-control study:
## use the following parameters
# outcome.type<-"binary"
# casecontrol<-TRUE
# prev<-0.12 ## prevalence of the disease (case) in general population; non-zero for a case-control study
# boot<-200  ## number of bootstrap samples
```


### Calculate the coefficients for each paths
```{r}
(results.tmp<-CensoredMediator(data,outcome.type,casecontrol,prev,printout=TRUE))

(IE.results<-results.tmp[[3]])

### Calculate the indirect effect (IE) and percentage mediated (PM)
### and their corresponding 95% confidence intervals using bootstrapping
### May take hours or days to run depending on the number of bootstrap samples
(results.boot<-mybootstrap(data,IE.results,boot,outcome.type,casecontrol,prev,printout=FALSE))

```

